---
title: "Group Project1_Monte Carlo Simulation Design."
author: "hw2849"
date: "02/15/2022 "
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(survival)
library(coxed)
library(nph)
library(tidyverse)
library(pwr)
library(simsurv)
library(flexsurv)
```


## Topic 2: Design a simulation study to assess the three hypothsis testings 

Design a set of distributions/models under both proportional-hazard and non-proportional-hazard assumptions, and carry out a simulation study to compare performance of those hypothesis tests in those models. Based on your numerical investigations, write a practical recommendation for general users to choose a suitable testing tool.  


## Scenario1: 
### Non-proportional

Under non-proportional test assumption, we generate data set of a sample size n = 1000, from two different distributions, exponential distribution (p = 500) and Weibull distribution (p = 500). We have a time variable, treatment and control group, and an event variable indicating censoring. 

```{r, non-proportional data simulation}
# sampling survival times from exponential distribution & Weibull distribution
set.seed(8160)
N = 500
treatment = rbinom(N, 1, 0.5) #trt or ctl

prop_data = data.frame(
  eventime = c(rexp(N, 0.1), rweibull(N, 0.6, 1)), 
  treatment
) %>% 
  mutate(
    status = case_when(eventime >= 4 ~ 0, 
                      eventime < 4 ~ 1)
  )

plot(survfit(Surv(eventime, status) ~ treatment, data = prop_data), 
    main = "Non-Proportional model",
     xlab = "Time", 
     ylab = "Overall survival probability")


## Non-proportional: log-rank and weighted log-rank tests 

logrank.test(prop_data$eventime, prop_data$status, prop_data$treatment, rho = 0, gamma = 0)
logrank.test(prop_data$eventime, prop_data$status, prop_data$treatment, rho = 1, gamma = 0)
logrank.test(prop_data$eventime, prop_data$status, prop_data$treatment, rho = 0, gamma = 1)
logrank.maxtest(prop_data$eventime, prop_data$status, prop_data$treatment)

## power test
ctl = prop_data %>% 
  filter(treatment == 0)
trt = prop_data %>% 
  filter(treatment == 1)
mu1 = mean(prop_data$eventime)
mu2 = mean(prop_data$eventime)
Var = var(prop_data$eventime)
H = (mu1- mu2)/Var
pwr = pwr.t2n.test(n1 = nrow(ctl), n2 = nrow(trt), d = H, sig.level = 0.05)
type2e = 1 - pwr$power
```


**Decisions**

*Log-rank test:* The Chi-Squared test statistic is 0.6 with 1 degree of freedom and the corresponding p-value is 0.5. Since this p-value is great than .05, we fail to reject the null hypothesis.

*Weighted log-rank test for an early effect:*
The Chi-Squared test statistic is 0.7 with 1 degree of freedom and the corresponding p-value is 0.4. Since this p-value is greater than 0.05, we fail to reject the null hypothesis.

*Weighted log-rank test for a late effect:*
The Chi-Squared test statistic is 0.2 with 1 degree of freedom and the corresponding p-value is 0.7. Since this p-value is greater than 0.05, we fail to reject the null hypothesis.

Therefore, when the hazard model simulated under non-proportional assumption with exponential and Weibull distribution, all tests failed to reject the null hypothesis. The type I errors are the p-values of each test. The type II error is 0.95.

## Scenario 2

### Proportional-censored

Under proportional assumption, we generate survival data with `sim.survdata` with variables `group`, `time`, and `fail`. Similarly, we simulated with a sample size of 1000, and tested `censored` and `uncensored`. 

* group: 1 = treatment, 0 = control

```{r, proportional simulation}
set.seed(8160)

## simulating data, censor.cond = TRUE
sim_data = sim.survdata(N = 500, T = 100, xvars = 1, censor = 0.2, censor.cond = TRUE, num.data.frames = 2) 

## data frame
prop_df = data.frame( #trt = 1, ctl = 0
  time_1 = sim_data[[1]]$data$y,
  time_0 = sim_data[[2]]$data$y,
  fail_1 = sim_data[[1]]$data$failed,
  fail_0 = sim_data[[2]]$data$failed
) %>% 
  pivot_longer( # make df readable
    time_1:fail_0,
    names_to = c(".value", "group"),
    names_sep = "_"
  ) %>% 
  mutate(group = as.numeric(group))

## plot survival curves for each group
plot(survfit(Surv(time, fail) ~ group, data = prop_df), 
     main = "Proportional model",
     xlab = "Time", 
     ylab = "Overall survival probability")

## Proportional: log-rank tests
logrank.test(prop_df$time, prop_df$fail, prop_df$group, rho = 0, gamma = 0)
logrank.test(prop_df$time, prop_df$fail, prop_df$group, rho = 1, gamma = 0)
logrank.test(prop_df$time, prop_df$fail, prop_df$group, rho = 0, gamma = 1)
logrank.maxtest(prop_df$time, prop_df$fail, prop_df$group)


## power test
ctl = prop_df %>% 
  filter(group == 0)
trt = prop_df %>% 
  filter(group == 1)
mu1 = mean(ctl$time)
mu2 = mean(trt$time)
Var = var(prop_df$time)
H = (mu1 - mu2)/Var
pwr = pwr.t2n.test(n1 = nrow(ctl), n2 = nrow(trt), d = H, sig.level = 0.05)
type2e = 1 - pwr$power
```

When n = 1000, T = 100, censor = 0.2, censor.cond = TRUE: 

**Decisions** 

*Log-rank test:* The Chi-Squared test statistic is 3.2 with 1 degree of freedom and the corresponding p-value is 0.07. Since this p-value is greater than .05, we fail to reject the null hypothesis.

*Weighted log-rank test for an early effect:*
The Chi-Squared test statistic is 2.4 with 1 degree of freedom and the corresponding p-value is 0.1. Since this p-value is greater than 0.05, we fail to reject the null hypothesis.

*Weighted log-rank test for a late effect:*
The Chi-Squared test statistic is 2.6 with 1 degree of freedom and the corresponding p-value is 0.1. Since this p-value is less than 0.05, we fail to reject the null hypothesis.

The power of the test is 0.949. 

#### Proportional: right-censored 

```{r, proportional simulation}
set.seed(8160)

## simulating data, censor.cond = false
sim_data = sim.survdata(N = 500, T = 100, xvars = 1, censor = 0.2, censor.cond = FALSE, num.data.frames = 2) 

## data frame
prop_df = data.frame( #trt = 1, ctl = 0
  time_1 = sim_data[[1]]$data$y,
  time_0 = sim_data[[2]]$data$y,
  fail_1 = sim_data[[1]]$data$failed,
  fail_0 = sim_data[[2]]$data$failed
) %>% 
  pivot_longer( # make df readable
    time_1:fail_0,
    names_to = c(".value", "group"),
    names_sep = "_"
  ) %>% 
  mutate(group = as.numeric(group))

## plot survival curves for each group
plot(survfit(Surv(time, fail) ~ group, data = prop_df), 
     main = "Proportional model",
     xlab = "Time", 
     ylab = "Overall survival probability")


logrank.test(prop_df$time, prop_df$fail, prop_df$group, rho = 0, gamma = 0)
logrank.test(prop_df$time, prop_df$fail, prop_df$group, rho = 1, gamma = 0)
logrank.test(prop_df$time, prop_df$fail, prop_df$group, rho = 0, gamma = 1)
logrank.maxtest(prop_df$time, prop_df$fail, prop_df$group)


## power test
ctl = prop_df %>% 
  filter(group == 0)
trt = prop_df %>% 
  filter(group == 1)
mu1 = mean(ctl$time)
mu2 = mean(trt$time)
Var = var(prop_df$time)
H = (mu1 - mu2)/Var
pwr = pwr.t2n.test(n1 = nrow(ctl), n2 = nrow(trt), d = H, sig.level = 0.05)
type2e = 1 - pwr$power
```

When n = 1000, T = 100, censor = 0.2, censor.cond = FALSE (right-censored): 

**Decisions** 

*Log-rank test:* The Chi-Squared test statistic is 104 with 1 degree of freedom and the corresponding p-value is 0. Since this p-value is less than .05, we reject the null hypothesis.

*Weighted log-rank test for an early effect:*
The Chi-Squared test statistic is 90.7 with 1 degree of freedom and the corresponding p-value is 0. Since this p-value is less than 0.05, we reject the null hypothesis.

*Weighted log-rank test for a late effect:*
The Chi-Squared test statistic is 65.3 with 1 degree of freedom and the corresponding p-value is 0. Since this p-value is less than 0.05, we reject the null hypothesis.

The power of the test is 0.938.


## repeat test for different conditions
Now we want to repeat our tests with different conditions. 
* censor = 0.9 

```{r, proportional simulation}
set.seed(8160)
## simulating data, censor.cond = TRUE
sim_data = sim.survdata(N = 500, T = 100, xvars = 1, censor = 0.9, censor.cond = TRUE, num.data.frames = 2) 

## data frame
prop_df = data.frame( #trt = 1, ctl = 0
  time_1 = sim_data[[1]]$data$y,
  time_0 = sim_data[[2]]$data$y,
  fail_1 = sim_data[[1]]$data$failed,
  fail_0 = sim_data[[2]]$data$failed
) %>% 
  pivot_longer( # make df readable
    time_1:fail_0,
    names_to = c(".value", "group"),
    names_sep = "_"
  ) %>% 
  mutate(group = as.numeric(group))

## plot survival curves for each group
plot(survfit(Surv(time, fail) ~ group, data = prop_df), 
     main = "Proportional model",
     xlab = "Time", 
     ylab = "Overall survival probability")

## Proportional: log-rank tests
logrank.test(prop_df$time, prop_df$fail, prop_df$group, rho = 0, gamma = 0)
logrank.test(prop_df$time, prop_df$fail, prop_df$group, rho = 1, gamma = 0)
logrank.test(prop_df$time, prop_df$fail, prop_df$group, rho = 0, gamma = 1)
logrank.maxtest(prop_df$time, prop_df$fail, prop_df$group)


## power test
ctl = prop_df %>% 
  filter(group == 0)
trt = prop_df %>% 
  filter(group == 1)
mu1 = mean(ctl$time)
mu2 = mean(trt$time)
Var = var(prop_df$time)
H = (mu1 - mu2)/Var
pwr = pwr.t2n.test(n1 = nrow(ctl), n2 = nrow(trt), d = H, sig.level = 0.05)
type2e = 1 - pwr$power
```

When n = 1000, T = 100, censor = 0.9, censor.cond = TRUE: 

**Decisions** 

*Log-rank test:* The Chi-Squared test statistic is 0.2 with 1 degree of freedom and the corresponding p-value is 0.7 Since this p-value is greater than .05, we fail to reject the null hypothesis.

*Weighted log-rank test for an early effect:*
The Chi-Squared test statistic is 0.1 with 1 degree of freedom and the corresponding p-value is 0.8 Since this p-value is greater than 0.05, we fail to reject the null hypothesis.

*Weighted log-rank test for a late effect:*
The Chi-Squared test statistic is 2.3 with 1 degree of freedom and the corresponding p-value is 0.1. Since this p-value is less than 0.05, we fail to reject the null hypothesis.



```{r, proportional simulation}
set.seed(8160)

## simulating data, censor.cond = false
sim_data = sim.survdata(N = 500, T = 100, xvars = 1, censor = 0.9, censor.cond = FALSE, num.data.frames = 2) 

## data frame
prop_df = data.frame( #trt = 1, ctl = 0
  time_1 = sim_data[[1]]$data$y,
  time_0 = sim_data[[2]]$data$y,
  fail_1 = sim_data[[1]]$data$failed,
  fail_0 = sim_data[[2]]$data$failed
) %>% 
  pivot_longer( # make df readable
    time_1:fail_0,
    names_to = c(".value", "group"),
    names_sep = "_"
  ) %>% 
  mutate(group = as.numeric(group))

## plot survival curves for each group
plot(survfit(Surv(time, fail) ~ group, data = prop_df), 
     main = "Proportional model",
     xlab = "Time", 
     ylab = "Overall survival probability")


logrank.test(prop_df$time, prop_df$fail, prop_df$group, rho = 0, gamma = 0)
logrank.test(prop_df$time, prop_df$fail, prop_df$group, rho = 1, gamma = 0)
logrank.test(prop_df$time, prop_df$fail, prop_df$group, rho = 0, gamma = 1)
logrank.maxtest(prop_df$time, prop_df$fail, prop_df$group)


## power test
ctl = prop_df %>% 
  filter(group == 0)
trt = prop_df %>% 
  filter(group == 1)
mu1 = mean(ctl$time)
mu2 = mean(trt$time)
Var = var(prop_df$time)
H = (mu1 - mu2)/Var
pwr = pwr.t2n.test(n1 = nrow(ctl), n2 = nrow(trt), d = H, sig.level = 0.05)
type2e = 1 - pwr$power
```

When n = 1000, T = 100, censor = 0.9, censor.cond = FALSE (right-censored): 

**Decisions** 

*Log-rank test:* The Chi-Squared test statistic is 10.4 with 1 degree of freedom and the corresponding p-value is 0001. Since this p-value is less than .05, we reject the null hypothesis.

*Weighted log-rank test for an early effect:*
The Chi-Squared test statistic is 11.9 with 1 degree of freedom and the corresponding p-value is 0. Since this p-value is less than 0.05, we reject the null hypothesis.

*Weighted log-rank test for a late effect:*
The Chi-Squared test statistic is 0.2 with 1 degree of freedom and the corresponding p-value is 0.6. Since this p-value is greater than 0.05, we fail to reject the null hypothesis.

## Conclusion

Under proportional assumption: 

* when the sample size increase, the overall survival probability increases. 
* when censor rate = 0.2, and censoring is true, all three tests fail to reject the null hypothesis; censoring is false (right-censored), all three tests reject the null hypothesis.
* similarly, when censor rate = 0.9, and censoring is true, all three tests fail to reject the null hypothesis; censoring is false (right-censored), weighted log-rank test for a late effect fail to reject the null hypothesis, and two other tests reject. 
0.9 true fail, false reject

Under non-proportional assumption: 
* Variation of sample size does not have a significant change on results; 
* All tests failed to reject the null hypothesis.